\chapter{Conclusion}

In this thesis, we introduced a novel network-level gradient path design for object
detection models, incorporating both ‘top-down’ and ‘bottom-up’ pathways in the
network’s ‘neck’. This method is versatile, with applicability to both single-stage
and multi-stage models.
Upon evaluating the performance of the original models and their modified
counterparts using our proposed method, we arrived at the following
conclusions:


The application of our method to single-stage models results in a more
significant reduction in computational resources compared to its
application to multi-stage models.
• Despite its lower efficiency with multi-stage models, our proposed method
can still be beneficial when the demand for computational resources
outweighs the need for model precision.
• The modified two-stage model, while being 9.51% lighter than its original
version, does exhibit a decrease in performance. Specifically, its mAP at a
0.5 IoU threshold is at least 7.8% lower, and the average mAP across IoU
thresholds from 0.5 to 0.95 is at least 3.6% lower.
• The one-stage model, when modified with our proposed method, becomes
32.67% lighter than its original version with a small decrease in mAP by
1.6% at 0.5 IoU threshold and average mAP across IoU thresholds from
0.5 to 0.95.
• Interestingly, on a remote sensing images dataset, the performance of the
modified one-stage model improves by at least 6.3% mAP at a 0.5 IoU
threshold and by 2.4% average mAP across IoU thresholds from 0.5 to
0.95.
In future work, we intend to further investigate the performance differences that
we have noticed between the single-stage and two-stage models. A key question
that emerged from our study is why the two-stage model did not yield similar
results to the single-stage model. Unravelling this mystery will be a primary focus,


as it has the potential to yield a useful understanding of the fundamental
mechanics of these models and inspire future techniques for optimisation.
Additionally, we plan to work on enhancing the performance of multi-stage
models. While our proposed method showed promising results with single-stage
models, its effectiveness was less noticeable when used with multi-stage models.
We believe there is potential for improvement in this area.
Lastly, to ensure the generalizability of our findings, we intend to experiment with
a broader range of single-stage models, including various versions of YOLO. This
will help us determine whether the results obtained from the TPH-YOLOv5
model apply to a diverse range of models, thereby solidifying the robustness of
our proposed method.